---
title: "P32 Sox9 OE Processed Data Analysis"
author: "Emily Butka"
date: "2023-08-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages}
library(Seurat)
library(dplyr)
library(reshape2)
library(grid)
library(ggplot2)
```

This notebook contains code used to analyze two P32 datasets with progenitors or preadipocytes over-expressing Sox9 following alignment and quality control. Code for all panels from Figure 4 and Supplementary Figure 7 is outlined here.

## Load processed data

Raw data are processed in the preceding notebook. Here, we will use the processed data for further analysis. 

Processed Sox9 over-expressed progenitors (experiment from Figure 4A) can be found here: https://figshare.com/s/c241b71c3d1ddf8e488a
Processed Sox9 over-expressed preadipocytes (experiment from Figure 4G) can be found here: https://figshare.com/s/7ca3418390a66a69d2e0

```{r load-data}
progs <- readRDS("~/Downloads/p32_sox9_oe_progenitors_processed.rds")
preadips <- readRDS("~/Downloads/p32_sox9_oe_preadipocytes_processed.rds")
```

## Color palettes

```{r}
colpal12 <- c("#FFC33B", "#FF6E3A", "#E20134", "#9F0162", "#FFB2FD", "#00C2F9", "#008DF9", "#8400CD", "#00FCCF", "#FF5AAF", "#009F81", "gray")
```

```{r}
colpal6 <- c("#211E71", "#009F81", "#FF5AAF", "#741112", "#FFC02D", "gray")
```

```{r}
colpal2_pink_blue <- c("#E75F62", "#232380")
```

## Capybara with P21 new classification reference

### Run Capybara using custom reference as defined by cell types in P21 dataset

Code and workflow described at https://github.com/morris-lab/Capybara. We did not perform tissue-level classification (Step 1); rather, a high-resolution custom reference was generated and continuous identity measured (Step 2) followed by discrete cell type classification and multiple identity scoring (Step 3).

Custom code to extract hybrid IDs and mean p-values for assignments is as follows, using data frames `bin.count` and `perc.list` described in Capybara workflow:

```{r}
# hybrids
bin.count.rowsums <- apply(bin.count, 1, sum)
rownames(bin.count)[which(bin.count.rowsums > 1)]
bin.count.greaterthan1 <- apply(bin.count[which(bin.count.rowsums > 1), ], c(1, 2), function(x) {x > 0})
multi.celltypes <- apply(bin.count.greaterthan1, 1, function(x) {print(colnames(bin.count.greaterthan1)[x])})

hybrids <- c()
for(i in 1:length(multi.celltypes)) {
  #print(multi.celltypes[i])
  hybrids <- c(hybrids, paste(multi.celltypes[[i]], collapse = "-"))
}

hybrids <- gsub("frxn_cell.type_", "", hybrids)

# p-values for classifications
pval_avg_mat <- data.frame()
for(i in 1:length(perc.list)) {
  pval_avg_mat <- rbind(pval_avg_mat, apply(perc.list[[i]][[1]], 2, mean))
  rownames(pval_avg_mat)[i] <- names(perc.list[[i]])
}
colnames(pval_avg_mat) <- names(apply(perc.list[[1]][[1]], 2, mean))
```

Classifications, hybrids, p-values, and identity fractions imported to progs and preadips Seurat objects as `capy_new_class`, `capy_new_class_hybrid`, `capy_pval_xxx`, and `capy_frxn_xxx`, respectively. This information is stored in `p32_naive_processed.rds` on Figshare.

### Figures and visualizations of Capybara data

```{r}
g_legend <- function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}
```

```{r}
ct_capy_prog <- data.frame(table(progs$celltag, progs$capy_new_class_short))
ct_capy_prog$Var2 <- factor(ct_capy_prog$Var2, levels = c("Prog", "Trans prog", "Imm preadip", "Comm preadip", "Multi ID"))
p4el <- ggplot(ct_capy_prog[ct_capy_prog$Var1 == "None", ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + theme_void() + ggtitle("Host\n(n = 11,502)") + theme(legend.position="bottom", plot.margin = margin(0, 0, 0, 0), legend.margin = margin(0, 0, 0, 0)) + labs(fill = "Cell type") + scale_fill_manual(values = colpal6[1:5]) + theme(plot.title = element_text(hjust = 0.5))
p4er <- ggplot(ct_capy_prog[ct_capy_prog$Var1 == "OE", ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + theme_void() + ggtitle("Sox9 OE\n(n = 2,467)") + theme(legend.position = "none") + scale_fill_manual(values =  colpal6[1:5]) + theme(plot.title = element_text(hjust = 0.5))
l4d <- g_legend(p4el)
#p5dl + p5dr
p4e <- grid.arrange(arrangeGrob(p4el + theme(legend.position="none"), p4er + theme(legend.position="none"), nrow=1), l4d, nrow=2, heights=unit(c(1.7, 0.3), c("in", "in")))
p4e
```

```{r}
ct_capy_preadip <- data.frame(table(preadips$celltag, preadips$capy_new_class_short))
ct_capy_preadip$Var2 <- factor(ct_capy_preadip$Var2, levels = c("Prog", "Trans prog", "Imm preadip", "Comm preadip", "Multi ID", "Unknown"))
p4hl <- ggplot(ct_capy_preadip[ct_capy_preadip$Var1 == "Ctrl", ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + ggtitle("Control Preadipocytes\n(n = 365)")+ theme_void() + theme(legend.position="bottom", plot.title = element_text(hjust = 0.5)) + labs(fill = "Cell type") + scale_fill_manual(values = colpal6)  #+ coord_flip()
p4hm <- ggplot(ct_capy_preadip[ct_capy_preadip$Var1 == "Dpp4", ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + theme_void() + ggtitle("Control Progenitors\n(n = 266)") + scale_fill_manual(values = colpal6) + theme(plot.title = element_text(hjust = 0.5)) #+ coord_flip()
p4hr <- ggplot(ct_capy_preadip[ct_capy_preadip$Var1 == "Sox9_OE", ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + theme_void() + ggtitle("Sox OE Preadipocytes\n(n = 538)") + scale_fill_manual(values = colpal6) + theme(plot.title = element_text(hjust = 0.5))#+ coord_flip()
l4h <- g_legend(p4hl)
p4h <- grid.arrange(p4hl + theme(legend.position="none"), p4hr + theme(legend.position="none"), p4hm + theme(legend.position="none"), nrow=1)
p4h
```

### Compare Capy results to P32 naive data

Processed and analyzed P32 naive data can be found on Figshare:

https://figshare.com/s/b75e3e90389bf38c5524

```{r load-data}
p32 <- readRDS("~/Downloads/p32_naive_processed.rds")
```

```{r}
ct_capy_naive <- data.frame(table(p32$celltag, p32$capy_new_class_short))
```

```{r}
ct_capy_prog$SoxOE <- "Yes"
ct_capy_naive$SoxOE <- "No"
ct_capy_prog$fraction <- ct_capy_prog$Freq / rep(table(progs$celltag), length(unique(ct_capy_prog$Var2)))
ct_capy_naive$fraction <- ct_capy_naive$Freq / rep(table(p32$celltag), length(unique(ct_capy_naive$Var2)))
ct_capy_prog_merge <- rbind(ct_capy_prog, ct_capy_naive)
```

```{r}
ct_capy_prog_merge$Var2 <- factor(ct_capy_prog_merge$Var2, levels = c("Prog", "Trans prog", "Imm preadip", "Comm preadip", "Multi ID", "Unknown"))
```

```{r}
p4f <- ggplot(ct_capy_prog_merge[ct_capy_prog_merge$Var1 %in% c("OE", "v2") & ct_capy_prog_merge$Var2 != "Unknown", ], aes(x = Var2, y = fraction, fill = SoxOE)) + geom_bar(stat = "identity", position = "dodge", color = "black") + theme_classic() + labs(x = "Cell type") + scale_fill_manual(values = c("#1C75BC", "#FFC33D"))
p4f
```

```{r}
ct_capy_preadip$SoxOE <- "Yes"
ct_capy_preadip$fraction <- ct_capy_preadip$Freq / rep(table(preadips$celltag), length(unique(ct_capy_preadip$Var2)))
ct_capy_preadip_merge <- rbind(ct_capy_preadip, ct_capy_naive)
```

```{r}
ct_capy_preadip_merge$Var2 <- factor(ct_capy_preadip_merge$Var2, levels = c("Prog", "Trans prog", "Imm preadip", "Comm preadip", "Multi ID", "Unknown"))
```

```{r}
p4i <- ggplot(ct_capy_preadip_merge[ct_capy_preadip_merge$Var1 %in% c("Sox9_OE", "v1") & ct_capy_preadip_merge$Var2 != "Unknown", ], aes(x = Var2, y = fraction, fill = SoxOE)) + geom_bar(stat = "identity", position = "dodge", color = "black") + theme_classic() + labs(x = "Cell type") + scale_fill_manual(values = c("#1C75BC", "#FFC33D"))
p4i
```

### Project progs and preadips datasets onto P32 naive embedding

```{r}
p32_model <- RunUMAP(p32, dims = 1:10, return.model = TRUE)
```

```{r}
p32.sox.prog.anchors <- FindTransferAnchors(reference = p32_model, query = progs,
    dims = 1:30, reference.reduction = "pca")
sox_prog_query <- MapQuery(anchorset = p32.sox.prog.anchors, reference = p32_model, query = progs,
    refdata = list(celltype = "capy_new_class"), reference.reduction = "pca", reduction.model = "umap")
```

```{r}
p32.sox.preadip.anchors <- FindTransferAnchors(reference = p32_model, query = preadips,
    dims = 1:30, reference.reduction = "pca")
sox_preadip_query <- MapQuery(anchorset = p32.sox.preadip.anchors, reference = p32_model, query = preadips,
    refdata = list(celltype = "capy_new_class"), reference.reduction = "pca", reduction.model = "umap")
```

Re-orient embedding if necessary

```{r}
p32_model@reductions$umap_mod <- p32_model@reductions$umap
p32_model@reductions$umap_mod@cell.embeddings[, "UMAP_2"] <- -1 * p32_model@reductions$umap_mod@cell.embeddings[, "UMAP_2"]
sox_prog_query@reductions$ref.umap_mod <- sox_prog_query@reductions$ref.umap
#colnames(sox_prog_query@reductions$ref.umapud@cell.embeddings) <- c("UMAP_1", "UMAP_2")
sox_prog_query@reductions$ref.umap_mod@cell.embeddings[, "refUMAP_2"] <- -1 * sox_prog_query@reductions$ref.umap_mod@cell.embeddings[, "refUMAP_2"]
sox_preadip_query@reductions$ref.umap_mod <- sox_preadip_query@reductions$ref.umap
#colnames(sox_preadip_query@reductions$ref.umapud@cell.embeddings) <- c("UMAP_1", "UMAP_2")
sox_preadip_query@reductions$ref.umap_mod@cell.embeddings[, "refUMAP_2"] <- -1 * sox_preadip_query@reductions$ref.umap_mod@cell.embeddings[, "refUMAP_2"]
```

```{r}
s7al <- DimPlot(p32_model, reduction = "umap_mod", group.by = "capy_new_class", cols = colpal6) + ggtitle("Reference annotations") + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), axis.title = element_blank(), text = element_text(size = 8), plot.margin = margin(0, 0, 0, 0), legend.position = "bottom")
s7am <- DimPlot(sox_prog_query, reduction = "ref.umap_mod", group.by = "predicted.celltype", cols = colpal6) + NoLegend() + ggtitle("Sox9 OE Progenitors,\nCellTagged controls, and host cells") + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), axis.title = element_blank(), text = element_text(size = 8), plot.margin = margin(0, 0, 0, 0), legend.position = "bottom")
s7ar <- DimPlot(sox_preadip_query, reduction = "ref.umap_mod", group.by = "predicted.celltype", cols = colpal6) + NoLegend() + ggtitle("Sox9 OE Preadipocytes,\nCellTagged controls, and host cells") + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), axis.title = element_blank(), text = element_text(size = 8), plot.margin = margin(0, 0, 0, 0), legend.position = "bottom")
ls7a <- g_legend(s7al)

s7a <- grid.arrange(arrangeGrob(s7al + theme(legend.position="none"), s7am + theme(legend.position="none"), s7ar + theme(legend.position = "none"), nrow=1), ls7a, nrow=2, heights=unit(c(1.7, 0.3), c("in", "in")))
s7a
```

```{r}
s7b1 <- ggplot(as.data.frame(sox_prog_query@reductions$ref.umap@cell.embeddings[sox_prog_query$celltag == "OE", ]), aes(x = refUMAP_1, y = -1 * refUMAP_2)) +
  stat_density_2d(aes(fill = stat(density)), geom="raster", contour = FALSE) +
  scale_fill_gradientn(colors = c("white", magma(5)[2:5])) +
  geom_point(alpha = 0.2, size = 0.2, color = "#FCFDBF") + theme_classic() + theme(axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank(), plot.margin = unit(c(0,0,0,0), "cm"), legend.position = "bottom")
s7b2 <- ggplot(as.data.frame(sox_prog_query@reductions$ref.umap@cell.embeddings[sox_prog_query$celltag == "None", ]), aes(x = refUMAP_1, y = -1 * refUMAP_2)) +
  stat_density_2d(aes(fill = stat(density)), geom="raster", contour = FALSE) +
  scale_fill_gradientn(colors = c("white", magma(5)[2:5])) +
  geom_point(alpha = 0.2, size = 0.2, color = "#FCFDBF") + theme_classic() + theme(axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank(), plot.margin = unit(c(0,0,0,0), "cm"), legend.position = "none")
s7b1 
s7b2
```

```{r}
s7c1 <- ggplot(as.data.frame(sox_preadip_query@reductions$ref.umap@cell.embeddings[sox_preadip_query$celltag == "Sox9_OE",]), aes(x = refUMAP_1, y = -1 * refUMAP_2)) +
  stat_density_2d(aes(fill = stat(density)), geom="raster", contour = FALSE) +
  scale_fill_gradientn(colors = c("white", magma(5)[2:5])) +
  geom_point(alpha = 0.2, size = 0.2, color = "#FCFDBF") + theme_classic() + theme(axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank(), plot.margin = unit(c(0,0,0,0), "cm"), legend.position = "bottom")
s7c2 <- ggplot(as.data.frame(sox_preadip_query@reductions$ref.umap@cell.embeddings[sox_preadip_query$celltag == "Dpp4",]), aes(x = refUMAP_1, y = -1 * refUMAP_2)) +
  stat_density_2d(aes(fill = stat(density)), geom="raster", contour = FALSE) +
  scale_fill_gradientn(colors = c("white", magma(5)[2:5])) +
  geom_point(alpha = 0.2, size = 0.2, color = "#FCFDBF") + theme_classic() + theme(axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank(), plot.margin = unit(c(0,0,0,0), "cm"), legend.position = "none")
s7c3 <- ggplot(as.data.frame(sox_preadip_query@reductions$ref.umap@cell.embeddings[sox_preadip_query$celltag == "Ctrl",]), aes(x = refUMAP_1, y = -1 * refUMAP_2)) +
  stat_density_2d(aes(fill = stat(density)), geom="raster", contour = FALSE) +
  scale_fill_gradientn(colors = c("white", magma(5)[2:5])) +
  geom_point(alpha = 0.2, size = 0.2, color = "#FCFDBF") + theme_classic() + theme(axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank(), plot.margin = unit(c(0,0,0,0), "cm"), legend.position = "none")
s7c4 <- ggplot(as.data.frame(sox_preadip_query@reductions$ref.umap@cell.embeddings[is.na(sox_preadip_query$celltag),]), aes(x = refUMAP_1, y = -1 * refUMAP_2)) +
  stat_density_2d(aes(fill = stat(density)), geom="raster", contour = FALSE) +
  scale_fill_gradientn(colors = c("white", magma(5)[2:5])) +
  geom_point(alpha = 0.2, size = 0.2, color = "#FCFDBF") + theme_classic() + theme(axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank(), plot.margin = unit(c(0,0,0,0), "cm"), legend.position = "none")
s7c1
s7c2
s7c3
s7c4
```

Again, we have deposited our processed datasets on Figshare. They have been analyzed according to the code in this notebook: 

Processed Sox9 over-expressed progenitors (experiment from Figure 4A) can be found here: https://figshare.com/s/c241b71c3d1ddf8e488a
Processed Sox9 over-expressed preadipocytes (experiment from Figure 4G) can be found here: https://figshare.com/s/7ca3418390a66a69d2e0
