---
title: "P32 Naive Processed Data Analysis"
author: "Emily Butka"
date: "2023-07-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages}
library(Seurat)
library(ggplot2)
library(scclusteval)
library(grid)
library(gridExtra)
```


This notebook contains code used to integrate 4 biological replicates, quality control, and analyze P32 naive data following alignment. Code for all panels from Figure 2, Figure 3A-B, and Supplementary Figure 5 is outlined here.

## Load processed data

Raw data are processed in the preceeding notebook. Here, we will use the processed data for further analysis. This can be found at: https://figshare.com/s/b75e3e90389bf38c5524

```{r load-data}
p32 <- readRDS("~/Downloads/p32_naive_processed.rds")
```

## Color palettes

```{r}
colpal12 <- c("#FFC33B", "#FF6E3A", "#E20134", "#9F0162", "#FFB2FD", "#00C2F9", "#008DF9", "#8400CD", "#00FCCF", "#FF5AAF", "#009F81", "gray")
```

```{r}
colpal6 <- c("#211E71", "#009F81", "#FF5AAF", "#741112", "#FFC02D", "gray")
```

```{r}
colpal2_pink_blue <- c("#E75F62", "#232380")
```

## Prelim plots

```{r}
p2b <- DimPlot(p32, label = T, label.size = 7, raster = F) + NoLegend() + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), axis.title = element_blank(), text = element_text(size = 8, family='Helvetica'), plot.margin = margin(0, 0, 0, 0)) 
p2b
```

```{r}
p2c <- DimPlot(p32, cells.highlight = list("Progenitors" = colnames(p32)[which(p32$celltag == "v2")], "Preadipocytes" = colnames(p32)[which(p32$celltag == "v1")]), cols.highlight = c("#232380", "#E75F62"), sizes.highlight = 0.5) + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), axis.title = element_blank(), text = element_text(size = 8), plot.margin = margin(0, 0, 0, 0), legend.position = "bottom")
p2c
```

```{r}
p3al <- FeaturePlot(p32, features = "Sox9")
p3ar <- FeaturePlot(p32, features = "Pparg")
p3al
p3ar
```

```{r}
p3b <- VlnPlot(p32, features = "Sox9", group.by = "celltag", pt.size = 0) + scale_fill_manual(values = c("gray", colpal2_pink_blue)) + theme(plot.title = element_blank()) + scale_x_discrete(labels = c("None", "Preadipocyte", "Progenitor")) + NoLegend()
p3b
```

```{r}
s5a <- FeaturePlot(p32, features = "Cebpa")
s5a
```

## Capybara with P21 new classification reference

### Run Capybara using custom reference as defined by cell types in P21 dataset

Code and workflow described at https://github.com/morris-lab/Capybara. We did not perform tissue-level classification (Step 1); rather, a high-resolution custom reference was generated and continuous identity measured (Step 2) followed by discrete cell type classification and multiple identity scoring (Step 3).

Custom code to extract hybrid IDs and mean p-values for assignments is as follows, using data frames `bin.count` and `perc.list` described in Capybara workflow:

```{r}
# hybrids
bin.count.rowsums <- apply(bin.count, 1, sum)
rownames(bin.count)[which(bin.count.rowsums > 1)]
bin.count.greaterthan1 <- apply(bin.count[which(bin.count.rowsums > 1), ], c(1, 2), function(x) {x > 0})
multi.celltypes <- apply(bin.count.greaterthan1, 1, function(x) {print(colnames(bin.count.greaterthan1)[x])})

hybrids <- c()
for(i in 1:length(multi.celltypes)) {
  #print(multi.celltypes[i])
  hybrids <- c(hybrids, paste(multi.celltypes[[i]], collapse = "-"))
}

hybrids <- gsub("frxn_cell.type_", "", hybrids)

# p-values for classifications
pval_avg_mat <- data.frame()
for(i in 1:length(perc.list)) {
  pval_avg_mat <- rbind(pval_avg_mat, apply(perc.list[[i]][[1]], 2, mean))
  rownames(pval_avg_mat)[i] <- names(perc.list[[i]])
}
colnames(pval_avg_mat) <- names(apply(perc.list[[1]][[1]], 2, mean))
```

Classifications, hybrids, p-values, and identity fractions imported to p32 Seurat object as `capy_new_class`, `capy_new_class_hybrid`, `capy_pval_xxx`, and `capy_frxn_xxx`, respectively. This information is stored in `p32_naive_processed.rds` on Figshare.

### Figures and visualizations of Capybara data

```{r}
p32$capy_new_class_short <- gsub("_", " ", p32$capy_new_class_short)
p32$capy_new_class_short <- factor(p32$capy_new_class_short, levels = c("Prog", "Trans int prog", "Imm preadip", "Comm preadip", "Multi ID", "Unknown"))
```

```{r}
g_legend <- function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}
```

```{r}
p2d <- DimPlot(p32, group.by = "capy_new_class_short", cols = colpal6) + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), axis.title = element_blank(), plot.title = element_blank(), plot.margin = margin(0, 0, 0, 0), legend.position = "bottom")
l2d <- g_legend(p2d)
p2d + NoLegend()
p2d <- grid.arrange(p2d + theme(legend.position = "none"), l2d, nrow = 2, heights = unit(c(2, 1), c("in", "in")))
p2d
```

```{r}
ctcapy <- data.frame(table(p32$celltag, p32$capy_new_class_short))
p2el <- ggplot(ctcapy[ctcapy$Var1 == "v2", ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + theme_void() + ggtitle("Dpp4 CellTagged cells") + theme(legend.position="none", text = element_text(size = 8), plot.title = element_text(hjust = 0.5)) + labs(fill = "Cell type") + scale_fill_manual(name = "Cell type", values = colpal6)
p2er <- ggplot(ctcapy[ctcapy$Var1 == "v1", ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + theme_void() + ggtitle("Cd9 CellTagged cells") + scale_fill_manual(name = "Cell type", values = colpal6) + theme(text = element_text(size = 8), legend.position = "none", plot.title = element_text(hjust = 0.5))
p2e <- p2el + p2er
p2e
```

```{r}
# Mark cells that have GFP > 0 OR a CellTag as "TRANSPLANT"
p32$is_transplant <- TRUE
p32$is_transplant[which(p32@assays$RNA@counts["GFP.CDS", ] == 0 & p32$celltag == "none")] <- FALSE
#table(p32$is_transplant)
```

```{r}
umap.embedding <- p32@reductions$umap@cell.embeddings
transplant.embedding <- as.data.frame(umap.embedding[which(p32$is_transplant), ])
```

```{r}
s5b <- ggplot(transplant.embedding, aes(x = UMAP_1, y = UMAP_2)) +
  stat_density_2d(aes(fill = stat(density)), geom="raster", contour = FALSE) +
  #scale_fill_viridis(option = "H") +
  #scale_fill_gradient(low = "white", high = "darkgreen") + 
  scale_fill_gradientn(colors = c("white", magma(5)[2:5])) +
  geom_point(alpha = 0.2, size = 0.2, color = "#FCFDBF") + theme_classic() + theme(axis.line = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), axis.title = element_blank(), plot.margin = unit(c(0,0,0,0), "cm"), legend.position = "bottom")
s5b + theme(legend.position = "bottom")
#ls5b <- g_legend(s5b)
#s5b <- grid.arrange(p3c + theme(legend.position = "none"), ls5b, nrow = 2, heights = unit(c(2, 1), c("in", "in")))
#p3cl
```

## Potential confounders

### Host vs. transplant

```{r}
# Mark cells that have GFP > 0 OR a CellTag as "TRANSPLANT"
p32$celltag_origin <- p32$celltag
p32$celltag_origin[which(p32@assays$RNA@counts["GFP.CDS", ] == 0 & p32$celltag == "none")] <- "None-Host"
p32$celltag_origin[which(p32@assays$RNA@counts["GFP.CDS", ] > 0 & p32$celltag == "none")] <- "None-Transplant"
```

#### Remove transplant cells, re-cluster, and compare clusters to original host-transplant co-embedding

```{r}
p32 <- FindNeighbors(p32, return.neighbor = T, k.param = 21)
```

```{r}
DefaultAssay(p32) <- "integrated"
```

```{r}
# New clusters and embedding of host cells only
p32_sub <- p32[, which(p32$celltag_origin == "None-Host")]
```

```{r}
# Clusters and embedding generated with host and transplant cells
p32_sub2 <- p32[, which(p32$celltag_origin == "None-Host")]
```

```{r}
p32_sub <- ScaleData(p32_sub)
p32_sub <- FindVariableFeatures(p32_sub)
p32_sub <- RunPCA(p32_sub)
p32_sub <- RunUMAP(p32_sub, dims = 1:10)
p32_sub <- FindNeighbors(p32_sub)
p32_sub <- FindClusters(p32_sub)
p32_sub <- FindNeighbors(p32_sub, return.neighbor = T, k.param = 21)
```

```{r}
DimPlot(p32_sub2, label = TRUE, label.size = 7) + NoLegend()
```

```{r}
DimPlot(p32_sub, label = TRUE, label.size = 7) + NoLegend()
```

```{r}
p32_sub2 <- FindNeighbors(p32_sub2, reduction = "umap", dims = 1:2, k.param = 21, return.neighbor = TRUE)
```

```{r}
nndistdf <- data.frame(p32_sub2@neighbors$integrated.nn@nn.dist[, 2:21], row.names = p32_sub2@neighbors$integrated.nn@cell.names)
```

```{r}
nndistdf_mean <- apply(nndistdf, 1, mean)
```

#### Now find distances and calculate mean for reclustered host data with same cells

```{r}
nncells_orig <- data.frame(p32_sub2@neighbors$integrated.nn@nn.idx[, 2:21], row.names = p32_sub2@neighbors$integrated.nn@cell.names)
```

```{r}
umap_new <- p32_sub@reductions$umap@cell.embeddings
```

```{r}
euclidean <- function(a, b) sqrt(sum((a - b)^2))
```

```{r}
nndistdf2 <- data.frame()
for(j in 1:dim(nncells_orig)[1]) {
  row <- c()
  for(i in 1:dim(nncells_orig)[2]) {
    row <- c(row, euclidean(umap_new[rownames(nncells_orig)[j], ], umap_new[nncells_orig[j, i], ]))
  }
  #print(row)
  nndistdf2 <- rbind(nndistdf2, row)
}
```

```{r}
nndistdf2_mean <- apply(nndistdf2, 1, mean)
```

```{r}
nncells_random <- data.frame()
for(j in 1:dim(nncells_orig)[1]) {
  row <- sample.int(dim(nncells_orig)[1], 20, replace = FALSE)
  nncells_random <- rbind(nncells_random, row)
}
rownames(nncells_random) <- p32_sub2@neighbors$integrated.nn@cell.names
```

```{r}
nndistdf_random <- data.frame()
for(j in 1:dim(nncells_orig)[1]) {
  row <- c()
  for(i in 1:dim(nncells_random)[2]) {
    row <- c(row, euclidean(umap_new[rownames(nncells_random)[j], ], umap_new[nncells_random[j, i], ]))
  }
  #print(row)
  nndistdf_random <- rbind(nndistdf_random, row)
}
```

```{r}
nndistdf_random_mean <- apply(nndistdf_random, 1, mean)
```

```{r}
nncells_random_in_clust <- data.frame()
for(j in 1:dim(nncells_orig)[1]) {
  clust <- p32_sub$seurat_clusters[rownames(nncells_orig)[j]]
  cells_in_clust <- c(1:dim(nncells_orig)[1])[which(p32_sub$seurat_clusters == clust)]
  row <- cells_in_clust[sample.int(length(cells_in_clust), 20, replace = FALSE)]
  nncells_random_in_clust <- rbind(nncells_random_in_clust, row)
}
rownames(nncells_random_in_clust) <- p32_sub2@neighbors$integrated.nn@cell.names
```

```{r}
nndistdf_random_in_clust <- data.frame()
for(j in 1:dim(nncells_orig)[1]) {
  row <- c()
  for(i in 1:dim(nncells_random_in_clust)[2]) {
    row <- c(row, euclidean(umap_new[rownames(nncells_random_in_clust)[j], ], umap_new[nncells_random_in_clust[j, i], ]))
  }
  #print(row)
  nndistdf_random_in_clust <- rbind(nndistdf_random_in_clust, row)
}
```

```{r}
nndistdf_random_in_clust_mean <- apply(nndistdf_random_in_clust, 1, mean)
```

```{r}
nndistdf2_own <- data.frame(p32_sub@neighbors$integrated.nn@nn.dist[, 2:21], row.names = p32_sub@neighbors$integrated.nn@cell.names)
```

```{r}
nndistdf2_own_mean <- apply(nndistdf2_own, 1, mean)
```

```{r}
dists <- data.frame("embedding" = c(rep("Original", dim(nncells_orig)[1]), rep("Reclustered", dim(nncells_orig)[1]), rep("Reclustered_own_20nn", dim(nncells_orig)[1]), rep("Random", dim(nncells_orig)[1]), rep("Random_in_clust", dim(nncells_orig)[1])),
                    "distance" = c(nndistdf_mean, nndistdf2_mean, nndistdf2_own_mean, nndistdf_random_mean, nndistdf_random_in_clust_mean))
```

```{r}
dists$embedding <- factor(dists$embedding, levels = c("Original", "Reclustered", "Reclustered_own_20nn", "Random", "Random_in_clust"))
```

```{r}
s5c <- ggplot(dists, aes(x = gsub("_", " ", embedding), y = distance)) + geom_boxplot() + theme_classic() + labs(x = "Embedding", y = "Distance")
s5c
ggsave("~/OneDrive/MorrisLab/Guillermo_project/manuscript/plots/updated_figs/recluster_boxplot.pdf", s5c, height = 2, width = 3, units = "in", dpi = 300)
```

### Louvain cluster

```{r}
s5d <- SilhouetteRainCloudPlot(CalculateSilhouette(adi_merged))
s5d
ggsave("~/OneDrive/MorrisLab/Guillermo_project/manuscript/plots/updated_figs/sillhouette.pdf", s5d, height = 2, width = 3, units = "in", dpi = 300)
```


### Biological replicate

#### Relative abundance of Capy cell types

```{r}
refqdfi <- data.frame(p32@meta.data[, c('celltag', 'capy_new_class', 'dataset')])
refqdfi <- data.frame(table(refqdfi$celltag, refqdfi$capy_new_class, refqdfi$dataset))
refqdfi$Var2 <- factor(refqdfi$Var2, levels = c("progenitor", "transitioning_progenitor", "immature_preadipocyte", "committed_preadipocyte", "Multi_ID"))
```

```{r}
colpal5 <- colpal6[1:5]
```

```{r}
p1 <- ggplot(refqdfi[which(refqdfi$Var1 == "none" & refqdfi$Var3 == "0706"), ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + coord_polar("y", start=0) + theme_void() + NoLegend() + scale_fill_manual(values = colpal5)
p2 <- ggplot(refqdfi[which(refqdfi$Var1 == "none" & refqdfi$Var3 == "0708"), ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + coord_polar("y", start=0) + theme_void() + NoLegend() + scale_fill_manual(values = colpal5)
p3 <- ggplot(refqdfi[which(refqdfi$Var1 == "none" & refqdfi$Var3 == "0722"), ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + coord_polar("y", start=0) + theme_void() + NoLegend() + scale_fill_manual(values = colpal5)
p4 <- ggplot(refqdfi[which(refqdfi$Var1 == "none" & refqdfi$Var3 == "1104"), ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + coord_polar("y", start=0) + theme_void() + NoLegend() + scale_fill_manual(values = colpal5)
p5 <- ggplot(refqdfi[which(refqdfi$Var1 == "v2" & refqdfi$Var3 == "0706"), ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + coord_polar("y", start=0) + theme_void() + NoLegend() + scale_fill_manual(values = colpal5)
p6 <- ggplot(refqdfi[which(refqdfi$Var1 == "v2" & refqdfi$Var3 == "0708"), ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + coord_polar("y", start=0) + theme_void() + NoLegend() + scale_fill_manual(values = colpal5)
p7 <- ggplot(refqdfi[which(refqdfi$Var1 == "v2" & refqdfi$Var3 == "0722"), ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + coord_polar("y", start=0) + theme_void() + NoLegend() + scale_fill_manual(values = colpal5)
p8 <- ggplot(refqdfi[which(refqdfi$Var1 == "v2" & refqdfi$Var3 == "1104"), ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + coord_polar("y", start=0) + theme_void() + NoLegend() + scale_fill_manual(values = colpal5)
p9 <- ggplot(refqdfi[which(refqdfi$Var1 == "v1" & refqdfi$Var3 == "0706"), ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + coord_polar("y", start=0) + theme_void() + NoLegend() + scale_fill_manual(values = colpal5)
p10 <- ggplot(refqdfi[which(refqdfi$Var1 == "v1" & refqdfi$Var3 == "0708"), ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + coord_polar("y", start=0) + theme_void() + NoLegend() + scale_fill_manual(values = colpal5)
p11 <- ggplot(refqdfi[which(refqdfi$Var1 == "v1" & refqdfi$Var3 == "0722"), ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + coord_polar("y", start=0) + theme_void() + NoLegend() + scale_fill_manual(values = colpal5)
p12 <- ggplot(refqdfi[which(refqdfi$Var1 == "v1" & refqdfi$Var3 == "1104"), ], aes(x="", y=Freq, fill=Var2)) + geom_bar(stat="identity", width=1, color="white") + coord_polar("y", start=0) + theme_void() + NoLegend() + scale_fill_manual(values = colpal5)
```

```{r}
t0 <- textGrob(" ", gp=gpar(fontfamily="Helvetica"))
t1 <- textGrob("No CT (Host)", gp=gpar(fontfamily="Helvetica"))
t2 <- textGrob("Dpp4 CT", gp=gpar(fontfamily="Helvetica"))
t3 <- textGrob("Cd9 CT", gp=gpar(fontfamily="Helvetica"))
t4 <- textGrob("Rep 1", gp=gpar(fontfamily="Helvetica"))
t5 <- textGrob("Rep 2", gp=gpar(fontfamily="Helvetica"))
t6 <- textGrob("Rep 3", gp=gpar(fontfamily="Helvetica"))
t7 <- textGrob("Rep 4", gp=gpar(fontfamily="Helvetica"))
```


```{r}
s5e <- grid.arrange(t0, t4, t5, t6, t7, t1, p1, p2, p3, p4, t2, p5, p6, p7, p8, t3, p9, p10, p11, p12, ncol = 5)
s5e
ggsave("~/OneDrive/MorrisLab/Guillermo_project/manuscript/plots/updated_figs/biological_replicates_p32.pdf", s5e, height = 2, width = 3, units = "in", dpi = 300)
```

```{r}
j1 <- reshape2::melt(PairWiseJaccardSets(p32$seurat_clusters, p32$dataset))
j2 <- reshape2::melt(PairWiseJaccardSets(p32$capy_new_class_short, p32$dataset))
s5h <- ggplot(j1, aes(x = factor(Var1), y = factor(Var2, levels = c("1104", "722", "708", "706")), fill = value)) + geom_tile(color = "white", lwd = 0.5, linetype = 1) + theme_classic() + theme(axis.line = element_blank(), axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + coord_fixed() + scale_fill_gradientn(colours = c("white", "darkred"), limits = c(0, 1)) + xlab("Louvain cluster") + ylab("Biological replicate") + theme(legend.position = "none")
s5i <- ggplot(j2, aes(x = factor(Var1, levels = c("Prog", "Trans prog", "Imm preadip", "Comm preadip", "Multi ID", "Unknown")), y = factor(Var2, levels = c("1104", "722", "708", "706")), fill = value)) + geom_tile(color = "white", lwd = 0.5, linetype = 1) + theme_classic() + theme(axis.line = element_blank(), axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + coord_fixed() + scale_fill_gradientn(colours = c("white", "darkred"), limits = c(0, 1)) + xlab("Cell type") + ylab("Biological replicate") + theme(legend.position = "none")
s5h
s5i
ggsave("~/OneDrive/MorrisLab/Guillermo_project/manuscript/plots/updated_figs/jaccard_biol_rep_clusters.pdf", s5h, height = 2, width = 3, units = "in", dpi = 300)
ggsave("~/OneDrive/MorrisLab/Guillermo_project/manuscript/plots/updated_figs/jaccard_biol_rep_celltype.pdf", s5i, height = 2, width = 3, units = "in", dpi = 300)
```


#### Jaccard analysis associating replicate with louvain cluster or cell type

Replicate x louvain cluster:

```{r}
jacc <- data.frame(PairWiseJaccardSets(p32$seurat_clusters, p32$dataset))
```

```{r}
num0722 <- table(p32$dataset)["0722"]
num1104 <- table(p32$dataset)["1104"]
num0706 <- table(p32$dataset)["0706"]
num0708 <- table(p32$dataset)["0708"]
p32$dataset_random <- sample(c(rep("0722", num0722), rep("1104", num1104), rep("0706", num0706), rep("0708", num0708)))
jacc_rand <- data.frame(PairWiseJaccardSets(p32$seurat_clusters, p32$dataset_random))
jacc_rand$rep <- 1
jacc_rand$cluster <- 0:13
```

```{r}
for(i in 2:1000) {
  p32$dataset_random <- sample(c(rep("0722", num0722), rep("1104", num1104), rep("0706", num0706), rep("0708", num0708)))
  jacc_rand_intermed <- data.frame(PairWiseJaccardSets(p32$seurat_clusters, p32$dataset_random))
  jacc_rand_intermed$rep <- i
  jacc_rand_intermed$cluster <- 0:13
  jacc_rand <- rbind(jacc_rand, jacc_rand_intermed)
}
```

```{r}
zscores <- c()
#mpzscores <- c()
for(i in 1:14) {
  z0722 <- (jacc$X0722[i] - mean(jacc_rand$X0722[jacc_rand$cluster == i - 1])) / sd(jacc_rand$X0722[jacc_rand$cluster == i - 1])
  z1104 <- (jacc$X1104[i] - mean(jacc_rand$X1104[jacc_rand$cluster == i - 1])) / sd(jacc_rand$X1104[jacc_rand$cluster == i - 1])
  z0706 <- (jacc$X0706[i] - mean(jacc_rand$X0706[jacc_rand$cluster == i - 1])) / sd(jacc_rand$X0706[jacc_rand$cluster == i - 1])
  z0708 <- (jacc$X0708[i] - mean(jacc_rand$X0708[jacc_rand$cluster == i - 1])) / sd(jacc_rand$X0708[jacc_rand$cluster == i - 1])
  zscores <- c(zscores, z0706, z0708, z0722, z1104)
}
```

```{r}
zdf <- data.frame(zscore = zscores, replicate = rep(c("Rep 1", "Rep 2", "Rep 3", "Rep 4"), 14), cluster = rep(0:13, each = 4))
```

```{r}
s5f <- ggplot(zdf, aes(x = cluster, y = replicate, fill = zscore)) + 
  geom_tile(color = "white",
            lwd = 0.5,
            linetype = 1) + geom_text(aes(label = round(zscore, digits = 1))) + theme_classic() + theme(axis.line = element_blank(), legend.position = "bottom", text = element_text(size = 8)) +
  scale_y_discrete("Replicate") +
  scale_x_continuous("Cluster", labels = as.character(seq(0,13)), breaks = seq(0,13)) +
  #scale_x_discrete("Cluster") + coord_fixed() +
  scale_fill_gradient2(
        low = "#053061", mid = "#F7F7F7", high = "#67001F", midpoint = 0)
s5f
```

Replicate x Capybara cell type:

```{r}
jacc <- data.frame(PairWiseJaccardSets(p32$capy_new_class_short, p32$dataset))
```

```{r}
num0722 <- table(p32$dataset)["0722"]
num1104 <- table(p32$dataset)["1104"]
num0706 <- table(p32$dataset)["0706"]
num0708 <- table(p32$dataset)["0708"]
p32$dataset_random <- sample(c(rep("0722", num0722), rep("1104", num1104), rep("0706", num0706), rep("0708", num0708)))
jacc_rand <- data.frame(PairWiseJaccardSets(p32$capy_new_class_short, p32$dataset_random))
jacc_rand$rep <- 1
jacc_rand$cell_type <- names(table(p32$capy_new_class_short))
```

```{r}
for(i in 2:1000) {
  p32$dataset_random <- sample(c(rep("0722", num0722), rep("1104", num1104), rep("0706", num0706), rep("0708", num0708)))
  jacc_rand_intermed <- data.frame(PairWiseJaccardSets(p32$capy_new_class_short, p32$dataset_random))
  jacc_rand_intermed$rep <- i
  jacc_rand_intermed$cell_type <- names(table(p32$capy_new_class_short))
  jacc_rand <- rbind(jacc_rand, jacc_rand_intermed)
}
```

```{r}
zscores <- c()
#mpzscores <- c()
for(i in names(table(p32$capy_new_class_short))) {
  z0722 <- (jacc[i, "X0722"] - mean(jacc_rand$X0722[jacc_rand$cell_type == i])) / sd(jacc_rand$X0722[jacc_rand$cell_type == i])
  z1104 <- (jacc[i, "X1104"] - mean(jacc_rand$X1104[jacc_rand$cell_type == i])) / sd(jacc_rand$X1104[jacc_rand$cell_type == i])
  z0706 <- (jacc[i, "X0706"] - mean(jacc_rand$X0706[jacc_rand$cell_type == i])) / sd(jacc_rand$X0706[jacc_rand$cell_type == i])
  z0708 <- (jacc[i, "X0708"] - mean(jacc_rand$X0708[jacc_rand$cell_type == i])) / sd(jacc_rand$X0708[jacc_rand$cell_type == i])
  zscores <- c(zscores, z0706, z0708, z0722, z1104)
}
```

```{r}
zdf <- data.frame(zscore = zscores, replicate = rep(c("Rep 1", "Rep 2", "Rep 3", "Rep 4"), 6), cell_type = rep(names(table(p32$capy_new_class_short)), each = 4))
zdf$cell_type <- factor(zdf$cell_type, levels = c("Prog", "Trans prog", "Imm preadip", "Comm preadip", "Multi ID", "Unknown"))
```

```{r}
s5g <- ggplot(zdf, aes(x = cell_type, y = replicate, fill = zscore)) + 
  geom_tile(color = "white",
            lwd = 0.5,
            linetype = 1) + geom_text(aes(label = round(zscore, digits = 1))) + theme_classic() + theme(axis.line = element_blank(), legend.position = "bottom", text = element_text(size = 8)) +
  scale_y_discrete("Replicate") +
  scale_x_discrete("Cell type") +
  #scale_x_discrete("Cluster") + coord_fixed() +
  scale_fill_gradient2(
        low = "#053061", mid = "#F7F7F7", high = "#67001F", midpoint = 0)
s5g
```

## Jaccard analysis associating cell type and louvain cluster

```{r}
p32$cp_ip <- ifelse(p32$capy_new_class_short == "Comm preadip", "comm_preadip", ifelse(p32$capy_new_class_short == "Prog", "prog", NA))
jacc <- data.frame(PairWiseJaccardSets(p32$seurat_clusters, p32$cp_ip))
```

```{r}
numip <- table(p32$capy_new_class_short)["Prog"]
numcp <- table(p32$capy_new_class_short)["Comm preadip"] 
numna <- sum(table(p32$capy_new_class_short)) - table(p32$capy_new_class_short)["Prog"] - table(p32$capy_new_class_short)["Comm preadip"] 
p32$cp_ip_random <- sample(c(rep("prog", numip), rep("comm_preadip", numcp), rep(NA, numna)))
jacc_rand <- data.frame(PairWiseJaccardSets(p32$seurat_clusters, p32$cp_ip_random))
jacc_rand$rep <- 1
jacc_rand$cluster <- 0:13
```

```{r}
for(i in 2:1000) {
  p32$cp_ip_random <- sample(c(rep("prog", numip), rep("comm_preadip", numcp), rep(NA, numna)))
  jacc_rand_intermed <- data.frame(PairWiseJaccardSets(p32$seurat_clusters, p32$cp_ip_random))
  jacc_rand_intermed$rep <- i
  jacc_rand_intermed$cluster <- 0:13
  jacc_rand <- rbind(jacc_rand, jacc_rand_intermed)
}
```

```{r}
mzscores <- c()
for(i in 1:14) {
  mzip <- 0.6745 * (jacc$int_prog[i] - median(jacc_rand$int_prog[jacc_rand$cluster == i - 1])) / mad(jacc_rand$int_prog[jacc_rand$cluster == i - 1])
  mzcp <- 0.6745 * (jacc$comm_preadip[i] - median(jacc_rand$comm_preadip[jacc_rand$cluster == i - 1])) / mad(jacc_rand$comm_preadip[jacc_rand$cluster == i - 1])
  mzscores <- c(mzscores, mzip, mzcp)
}
```

```{r}
zdf <- data.frame(mzscore = mzscores, cell_type = rep(c("int_prog", "comm_preadip"), 14), cluster = c(0, 0, 1, 1, 2, 2, 3, 3, 4, 4, 5, 5, 6, 6, 7, 7, 8, 8, 9, 9, 10, 10, 11, 11, 12, 12, 13, 13))
```

```{r}
p2f <- ggplot(zdf, aes(x = cluster, y = cell_type, fill = mzscore)) + 
  geom_tile(color = "white",
            lwd = 0.5,
            linetype = 1) + geom_text(aes(label = round(mzscore))) + theme_classic() + theme(axis.line = element_blank(), legend.position = "bottom", text = element_text(size = 8)) +
  scale_y_discrete("Cell type") +
  scale_x_continuous("Cluster", labels = as.character(seq(0,13)), breaks = seq(0,13)) +
  #scale_x_discrete("Cluster") + coord_fixed() +
  scale_fill_gradient2(
        low = "#053061", mid = "#F7F7F7", high = "#67001F", midpoint = 0)
p2f
```

## Permutation testing

Permutation testing was performed in Python to test whether louvain clusters are enriched for one CellTagged population or the other. Cluster and CellTag information was used; these data and code for the test are posted on GitHub.

Cells expressing the V1 CellTag were found to be enriched in clusters 1, 4, 5, 10, and 12. Cells expressing the V2 CellTag were found to be enriched in clusters 1, 3, 6, 9, 10, and 12. This information is stored in the P32 Seurat object as `enriched`.

```{r}
df2g <- data.frame(UMAP_1 = p32@reductions$umap@cell.embeddings[, 1], UMAP_2 = p32@reductions$umap@cell.embeddings[, 2], enriched = p32$enriched)
df2g <- rbind(df2g, data.frame(UMAP_1 = p32@reductions$umap@cell.embeddings[which(p32$enriched == "Both"), 1], UMAP_2 = p32@reductions$umap@cell.embeddings[which(p32$enriched == "Both"), 2], enriched = rep("V1", table(p32$enriched)["Both"])))
df2g$enriched <- gsub("Both", "V2", df2g$enriched)
df2g2 <- na.omit(df2g)
```

```{r}
p2g <- DimPlot(p32, label = T, label.size = 7, cols = c(rep("gray", 14), colpal2_pink_blue)) + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), axis.title = element_blank(), plot.margin = margin(0, 0, 0, 0)) + NoLegend() + geom_density_2d(data = df2g2, aes(x = UMAP_1, y = UMAP_2, color = enriched))
p2g
```

```{r}
s5h <- DimPlot(adi_merged, cells.highlight = list("Dpp4" = colnames(adi_merged)[which(adi_merged$enriched == "V2")], "Cd9" = colnames(adi_merged)[which(adi_merged$enriched == "V1")], "Both" = colnames(adi_merged)[which(adi_merged$enriched == "Both")]), cols.highlight = c("#232380", "#E75F62", "#1DD3B0")) + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), axis.title = element_blank(), text = element_text(size = 8), plot.margin = margin(0, 0, 0, 0), legend.position = "bottom")
s5h + theme(legend.position = "none")
ls5h <- g_legend(s5h)
s5hl <- grid.arrange(s5h + theme(legend.position = "none"), ls5h, nrow = 2, heights = unit(c(2, 1), c("in", "in")))
s5hl
```

```{r}
celltag_dist_df <- data.frame(table(p32$seurat_clusters, p32$celltag))
```

```{r}
celltag_dist_df$percent <- NA
celltagsums <- table(p32$celltag)
for(i in 1:dim(celltag_dist_df)[1]) {
  celltag_dist_df$percent[i] <- celltag_dist_df$Freq[i] / table(p32$celltag)[celltag_dist_df$Var2[i]]
}
```

```{r}
celltag_dist_df$Var1 <- factor(celltag_dist_df$Var1, levels = c(3, 6, 9, 1, 10, 12, 4, 5, 0, 2, 7, 8, 11, 13))
```

```{r}
s5i <- ggplot(celltag_dist_df[which(celltag_dist_df$Var2 != "none"), ], aes(x = Var1, y = percent, fill = Var2)) + geom_bar(stat = "identity", position = position_dodge(), color = "black") + theme_classic() + labs(fill = "CellTag", y = "Fraction", x = "Cluster") + scale_fill_manual(values =  colpal2_pink_blue, labels = c("Cd9", "Dpp4")) + theme(legend.position = "none")
s5i
```

Again, we have deposited our whole processed P21 dataset on Figshare. It has been analyzed according to the code in this notebook: 
https://figshare.com/s/b75e3e90389bf38c5524



## 2/3/24 plots

```{r}
s5a <- FeaturePlot(adi_merged, features = "Cebpa", order = T) + NoLegend() + theme(line = element_blank(),
                                                                axis.title = element_blank(),
                                                                axis.text = element_blank(),
                                                                plot.title = element_blank(),
                                                                plot.margin = margin(0, 0, 0, 0))
s5a
ggsave("~/OneDrive/MorrisLab/Guillermo_project/manuscript/plots/updated_figs/cebpa_p32.pdf", s5a, height = 1, width = 1.5, units = "in", dpi = 300)
```

```{r}
s5l <- VlnPlot(p32, features = "Dpp4", sort = TRUE, pt.size = 0) + NoLegend()
s5m <- VlnPlot(p32, features = "Cd9", sort = TRUE, pt.size = 0) + NoLegend()
s5l
s5m
ggsave("~/OneDrive/MorrisLab/Guillermo_project/manuscript/plots/updated_figs/dpp4_violin_p32.pdf", s5l, height = 2, width = 3, units = "in", dpi = 300)
ggsave("~/OneDrive/MorrisLab/Guillermo_project/manuscript/plots/updated_figs/cd9_violin_p32.pdf", s5m, height = 2, width = 3, units = "in", dpi = 300)
```

```{r}

```





