---
title: "P21 Processed Data Analysis"
author: "Emily Butka"
date: "2023-07-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load-packages}
library(Seurat)
library(dplyr)
library(reshape2)
#library(grid)
library(ggplot2)
library(RColorBrewer)
```

This notebook contains code used to analyze P21 data following alignment and quality control. Code for all panels from Figure 1 and Supplementary Figure 1 is outlined here.

## Load processed data

Raw data are processed in the preceeding notebook. Here, we will use the processed data for further analysis. This can be found at: https://figshare.com/s/396c5f4ebd3c23d20094

```{r load-data}
p21 <- readRDS("~/Downloads/p21_processed.rds")
```

## Color palettes

```{r}
colpal12 <- c("#FFC33B", "#FF6E3A", "#E20134", "#9F0162", "#FFB2FD", "#00C2F9", "#008DF9", "#8400CD", "#00FCCF", "#FF5AAF", "#009F81", "gray")
```

```{r}
colpal6 <- c("#211E71", "#009F81", "#FF5AAF", "#741112", "#FFC02D", "gray")
```

```{r}
capy_ref_colpal <- c("#756BB1", "#E6550D", "#DE2D26", "#3182BD", "#31A354", "#FFC02D")
```

## Preliminary Plot

```{r}
p1b <- DimPlot(p21, label = T, label.size = 7, raster = F) + NoLegend() + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), axis.title = element_blank(), text = element_text(size = 8, family='Helvetica'))
p1b
```

## Capybara with 5 published references

```{r}
all_capys <- rbind(data.frame(call = p21$capy_classification_merrick_etal), data.frame(call = p21$capy_phan), data.frame(call = p21$capy_jacob), data.frame(call = p21$capy_buechler_clustname), data.frame(call = p21$capy_burl))
all_capys$X <- NULL
all_capys$reference <- c(rep("Merrick", dim(p21)[2]), 
                         rep("Phan", dim(p21)[2]),
                         rep("Jacob", dim(p21)[2]),
                         rep("Buechler", dim(p21)[2]),
                         rep("Burl", dim(p21)[2]))
```

```{r}
dfplot <- data.frame(table(all_capys$call, all_capys$reference))
```

```{r}
ggplot(dfplot, aes(x = Var2, y = Freq / sum(Freq) * 5, fill = Var1)) + geom_col() + theme_classic()
```

### factor cell types

```{r}
# Collapse unassigned into unknown, remove unassigned
dfplot$Freq[dfplot$Var2 == "Merrick" & dfplot$Var1 == "Unknown"] <- dfplot$Freq[dfplot$Var2 == "Merrick" & dfplot$Var1 == "Unassigned"]
dfplot <- dfplot[which(dfplot$Var1 != "Unassigned"), ]
```

```{r}
dfplot$Var1 <- gsub("_", " ", dfplot$Var1)
dfplot$Var1 <- gsub(" cd142", "", dfplot$Var1)
dfplot$Var1 <- gsub("group", "Group", dfplot$Var1)
dfplot$Var1 <- gsub("\\.", " ", dfplot$Var1)
```

```{r}
ggplot(dfplot, aes(x = Var2, y = Freq / sum(Freq) * 5, fill = Var1)) + geom_col() + theme_classic() + xlab("Reference") + ylab("Relative Abundance")
```

```{r}
dfplot$Var1 <- factor(dfplot$Var1, levels = c("Ccl19", "Col15a1", "Comp", "Cxcl12", "Fbln1", "Hhip", "Npnt", "Pi16", "ASC1", "ASC2", "FIB Inter1", "FIB Inter2", "FIB Inter3", "MUSCLE Late", "VESSEL LECs", "adipocyte", "committed preadipocyte", "endothelial", "Group 3", "Group 4", "Group 5", "Group 6", "interstitial progenitor", "neural crest", "smooth muscle", "fibroblast", "keratinocyte", "lymphocyte blood", "pericytes", "Unknown", "Multi ID"))
```

```{r}
s1b <- ggplot(dfplot, aes(x = Var2, y = Freq / sum(Freq) * 5, fill = Var1)) + geom_col() + theme_classic() + xlab("Reference") + ylab("Relative Abundance") + scale_fill_manual(name = "Cell type", values = c(brewer.pal(9, "Reds")[2:9], brewer.pal(3, "Blues")[2:3], brewer.pal(5, "Oranges"), brewer.pal(9, "Purples"), "#25006D", brewer.pal(5, "Greens")[2:5], "gray", "yellow")) #+ theme(legend.position = "none")
s1b
```

## Capybara with consolidated reference

```{r}
dfplot <- data.frame(table(p21$capy_super_ref))
dfplot$ident <- "P21"
ggplot(dfplot, aes(x = ident, y = Freq / sum(Freq), fill = Var1)) + geom_col() + theme_classic() #+ theme(legend.position = "bottom")
```

```{r}
identical(colnames(p21), super_ref_class$barcode)
p21$capy_super_ref <- super_ref_class$call
```

```{r}
super_ref_class$origin <- super_ref_class$call
super_ref_class$origin[super_ref_class$origin %in% c("Ccl19", "Col15a1", "Comp", "Cxcl12", "Fbln1", "Hhip", "Npnt", "Pi16")] <- "Buechler"
super_ref_class$origin[super_ref_class$origin %in% c("ASC1", "ASC2")] <- "Burl"
super_ref_class$origin[super_ref_class$origin %in% c("FIB Inter1", "FIB Inter2", "FIB Inter3", "MUSCLE.Late", "VESSEL LECs")] <- "Jacob"
super_ref_class$origin[super_ref_class$origin %in% c("adipocyte", "committed_preadipocyte", "endothelial", "group_3_cd142", "group_4", "group_5", "group_6", "interstitial_progenitor", "neural_crest", "smooth_muscle")] <- "Merrick"
super_ref_class$origin[super_ref_class$origin %in% c("fibroblast", "keratinocyte", "lymphocyte blood", "pericytes")] <- "Phan"
```

```{r}
dfplot <- data.frame(table(super_ref_class$origin))
dfplot$ident <- "P21 Skin"
dfplot$Var1 <- gsub("_", " ", dfplot$Var1)
dfplot$Var1 <- factor(dfplot$Var1, levels = c("Merrick", "Jacob", "Burl", "Buechler", "Phan", "Multi ID"))
```

```{r}
p1c <- ggplot(dfplot, aes(x = ident, y = Freq / sum(Freq), fill = Var1)) + geom_col() + theme_classic() + scale_fill_manual(name = "Dataset of origin of\nCapybara\nclassification", values = capy_ref_colpal) + theme(axis.title.x = element_blank(), axis.line.x = element_blank()) + ylab("Relative abundance")
p1c
```

### Nearest neighbors analysis

P1D

### Hybrids analysis

```{r}
super_ref_hybrid <- read.csv("~/OneDrive/MorrisLab/Guillermo_project/23_08_28_capy_ref_comparisons/23_11_03_super_ref_on_p21_hybrids.csv")
rownames(super_ref_hybrid) <- super_ref_hybrid$x
```

```{r}
celltypes <- c()
for(i in super_ref_hybrid$hybrid) {
  types <- strsplit(i, "-")
  if(!(types[[1]][1] %in% celltypes)) {
    celltypes <- c(celltypes, types[[1]][1])
  }
  if(!(types[[1]][2] %in% celltypes)) {
    celltypes <- c(celltypes, types[[1]][2])
  }
}
  
```

```{r}
hybrid_ids <- data.frame(matrix(rep(0, length(celltypes)^2), nrow = length(celltypes)))
colnames(hybrid_ids) <- celltypes
rownames(hybrid_ids) <- colnames(hybrid_ids)
```

```{r}
hybrids_table <- data.frame(table(super_ref_hybrid$hybrid))
```

```{r}
count <- 0
cellscount <- 0
for(i in 1:dim(hybrids_table)[1]) {
  types <- strsplit(as.character(hybrids_table$Var1[i]), "-")
  if(length(types[[1]]) == 2) {
    hybrid_ids[types[[1]][1], types[[1]][2]] <- hybrid_ids[types[[1]][1], types[[1]][2]] + hybrids_table$Freq[i]
  }
  else {
    count <- count + 1
    cellscount <- cellscount + hybrids_table$Freq[i]
  }
}
print(count)
print(cellscount)
hybrid_ids <- reshape2::melt(hybrid_ids)
hybrid_ids$variable2 <- rep(unique(hybrid_ids$variable), length(celltypes))

```

```{r}
ggplot(hybrid_ids, aes(x=variable, y=variable2, size=ifelse(value==0, NA,  value / sum(hybrid_ids$value)))) + geom_point(aes(color = variable)) +
  scale_size_area(name = "Fraction of hybrids", max_size=6) +
  guides(fill = FALSE, color = FALSE) +
  theme(legend.position = "bottom",
        axis.ticks = element_blank(), 
        axis.title = element_blank(), 
        legend.key=element_blank(),
        legend.box.spacing = unit(0, "in"),
        legend.margin = margin(0, 0, 0, 0),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line.x = element_line(color = "black", size = 1),
        axis.line.y = element_line(color = "black", size = 1), axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))


```

```{r}
hybrid_ids$variable_origin <- hybrid_ids$variable
hybrid_ids$variable_origin <- as.character(hybrid_ids$variable_origin)
hybrid_ids$variable_origin[hybrid_ids$variable_origin %in% c("Ccl19", "Col15a1", "Comp", "Cxcl12", "Fbln1", "Hhip", "Npnt", "Pi16")] <- "Buechler"
hybrid_ids$variable_origin[hybrid_ids$variable_origin %in% c("ASC1", "ASC2")] <- "Burl"
hybrid_ids$variable_origin[hybrid_ids$variable_origin %in% c("FIB Inter1", "FIB Inter2", "FIB Inter3", "MUSCLE.Late", "VESSEL LECs", "EPI.Periderm")] <- "Jacob"
hybrid_ids$variable_origin[hybrid_ids$variable_origin %in% c("adipocyte", "committed_preadipocyte", "endothelial", "group_3_cd142", "group_4", "group_5", "group_6", "interstitial_progenitor", "neural_crest", "smooth_muscle")] <- "Merrick"
hybrid_ids$variable_origin[hybrid_ids$variable_origin %in% c("fibroblast", "keratinocyte", "lymphocyte blood", "pericytes")] <- "Phan"
```

```{r}
hybrid_ids$variable_origin2 <- hybrid_ids$variable2
hybrid_ids$variable_origin2 <- as.character(hybrid_ids$variable_origin2)
hybrid_ids$variable_origin2[hybrid_ids$variable_origin2 %in% c("Ccl19", "Col15a1", "Comp", "Cxcl12", "Fbln1", "Hhip", "Npnt", "Pi16")] <- "Buechler"
hybrid_ids$variable_origin2[hybrid_ids$variable_origin2 %in% c("ASC1", "ASC2")] <- "Burl"
hybrid_ids$variable_origin2[hybrid_ids$variable_origin2 %in% c("FIB Inter1", "FIB Inter2", "FIB Inter3", "MUSCLE.Late", "VESSEL LECs", "EPI.Periderm")] <- "Jacob"
hybrid_ids$variable_origin2[hybrid_ids$variable_origin2 %in% c("adipocyte", "committed_preadipocyte", "endothelial", "group_3_cd142", "group_4", "group_5", "group_6", "interstitial_progenitor", "neural_crest", "smooth_muscle")] <- "Merrick"
hybrid_ids$variable_origin2[hybrid_ids$variable_origin2 %in% c("fibroblast", "keratinocyte", "lymphocyte blood", "pericytes")] <- "Phan"
```

```{r}
hybrid_ids_origin <- aggregate(value~variable_origin+variable_origin2,data=hybrid_ids,FUN=sum)
```

```{r}
hybrid_ids_origin$variable_origin <- factor(hybrid_ids_origin$variable_origin, levels = c("Merrick", "Jacob", "Burl", "Buechler", "Phan"))
hybrid_ids_origin$variable_origin2 <- factor(hybrid_ids_origin$variable_origin2, levels = c("Merrick", "Jacob", "Burl", "Buechler", "Phan"))
```

```{r}
s1c <- ggplot(hybrid_ids_origin, aes(x=variable_origin, y=variable_origin2, size=ifelse(value==0, NA,  value / sum(hybrid_ids_origin$value)))) + geom_point(aes(color = variable_origin)) +
  scale_size_area(name = "Fraction of hybrids", max_size=6) +
  guides(fill = "legend", color = FALSE) +
  theme(legend.position = "bottom",
        axis.ticks = element_blank(), 
        axis.title = element_blank(), 
        legend.key=element_blank(),
        legend.box.spacing = unit(0, "in"),
        legend.margin = margin(0, 0, 0, 0),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line.x = element_line(color = "black", size = 1),
        axis.line.y = element_line(color = "black", size = 1), axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) + scale_color_manual(values = capy_ref_colpal)
s1c
```

## Capybara with Merrick, et al reference

### Download and process Merrick, et al dataset from GEO to recapitulate cell types as defined by authors (GEO accession GSE128889; GSM3717977)

```{r}
Merrick_etal.data <- Read10X(data.dir = "~/Downloads/GSE128889/GSM3717977/") # this directory to contain 3 files: GSM3717977_SCmurinep12_barcodes.tsv.gz, GSM3717977_SCmurinep12_genes.tsv.gz, GSM3717977_SCmurinep12_matrix.mtx.gz
merrick_p12 <- CreateSeuratObject(counts = Merrick_etal.data, min.cells = 3, min.features = 200, project = "Seale_P12")
merrick_p12[["percent.mt"]] <- PercentageFeatureSet(object = merrick_p12, pattern = "^mt-")
merrick_p12 <- subset(x = merrick_p12, subset = nFeature_RNA > 200 & nFeature_RNA < 3600 & percent.mt < 7)
merrick_p12 <- NormalizeData(object = merrick_p12, normalization.method = "LogNormalize", scale.factor = 10000)
merrick_p12 <- FindVariableFeatures(object = merrick_p12, selection.method = "vst", nfeatures = 2000)
merrick_p12 <- ScaleData(object = merrick_p12, features = rownames(x = merrick_p12))
merrick_p12 <- RunPCA(object = merrick_p12, features = VariableFeatures(object = merrick_p12))
merrick_p12 <- JackStraw(object = merrick_p12, num.replicate = 100)
merrick_p12 <- ScoreJackStraw(object = merrick_p12, dims = 1:20)
JackStrawPlot(object = merrick_p12, dims = 1:20)
```

```{r}
ElbowPlot(object = merrick_p12)
```

```{r}
merrick_p12 <- FindNeighbors(object = merrick_p12, dims = 1:20)
merrick_p12 <- FindClusters(object = merrick_p12, resolution = 0.6)
merrick_p12 <- RunUMAP(object = merrick_p12, dims = 1:15)
```

```{r}
# find markers for every cluster compared to all remaining cells, report only the positive ones
merrick_p12.markers <- FindAllMarkers(object = merrick_p12, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
merrick_p12.markers %>% group_by(cluster) %>% top_n(n = 40, wt = avg_log2FC)
```

We have deposited our processed data from Merrick, et al here, for downstream analysis: https://figshare.com/s/224f9152394137f8c46d

```{r}
merrick_p12 <- readRDS("~/Downloads/merrick_etal_processed.rds")
```

### Merrick et al. prelim plots

```{r}
sm1a <- DimPlot(merrick_p12, label = T, label.size = 7) + NoLegend() + theme(axis.title = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), axis.text = element_blank())
sm1c
```

```{r}
sm1b <- VlnPlot(object = merrick_p12, features = c("Dpp4", "Wnt2", "Akr1c18", "Anxa3", "Sema3c", "Icam1", "Dlk1", "Fabp4", "Pparg", "Ggt5", "Cd9", "Egfl6", "Emb", "F3", "Fmo2", "Gdf10", "Ifitm1", "Thbs4", "Plxdc1", "Adipoq", "Cd36", "Plin1", "Pecam1", "Acta2", "Mpz"), stack = T, flip = T) + NoLegend()
sm1b
```

### Run Capybara using custom reference with cell types as defined above 

Code and workflow described at https://github.com/morris-lab/Capybara. We did not perform tissue-level classification (Step 1); rather, a high-resolution custom reference was generated and continuous identity measured (Step 2) followed by discrete cell type classification and multiple identity scoring (Step 3).

Custom code to extract hybrid IDs and mean p-values for assignments is as follows, using data frames `bin.count` and `perc.list` described in Capybara workflow:

```{r}
# hybrids
bin.count.rowsums <- apply(bin.count, 1, sum)
rownames(bin.count)[which(bin.count.rowsums > 1)]
bin.count.greaterthan1 <- apply(bin.count[which(bin.count.rowsums > 1), ], c(1, 2), function(x) {x > 0})
multi.celltypes <- apply(bin.count.greaterthan1, 1, function(x) {print(colnames(bin.count.greaterthan1)[x])})

hybrids <- c()
for(i in 1:length(multi.celltypes)) {
  #print(multi.celltypes[i])
  hybrids <- c(hybrids, paste(multi.celltypes[[i]], collapse = "-"))
}

hybrids <- gsub("frxn_cell.type_", "", hybrids)

# p-values for classifications
pval_avg_mat <- data.frame()
for(i in 1:length(perc.list)) {
  pval_avg_mat <- rbind(pval_avg_mat, apply(perc.list[[i]][[1]], 2, mean))
  rownames(pval_avg_mat)[i] <- names(perc.list[[i]])
}
colnames(pval_avg_mat) <- names(apply(perc.list[[1]][[1]], 2, mean))
```

Classifications, hybrids, p-values, and identity fractions imported to p21 Seurat object as `capy_classification_merrick_etal`, `capy_hybrid_merrick_etal`, `pval_cell.type_xxx_merrick`, and `frxn_cell.type_xxx_merrick`, respectively. This information is stored in `p21_processed.rds` on Figshare.

### Figures and visualizations of Capybara data

```{r}
p1el <- DimPlot(p21, group.by = "capy_classification_merrick_etal", pt.size = 1) + scale_color_manual(labels = c("Adipocyte", "Committed preadipocyte", "Endothelial", "Group 3", "Group 4", "Group 5", "Group 6", "Interstitial progenitor", "Multi ID", "Neural crest", "Smooth muscle", "Unassigned"), values = colpal12) + theme(plot.title = element_blank(), axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), legend.position = "bottom", text = element_text(size = 8), plot.margin = margin(0, 0, 0, 0), legend.margin = margin(0, 0, 0, 0))
p1el + NoLegend()
```

```{r}
skin_inguinal_celltypes <- data.frame(table(c(rep("P21 Skin", dim(p21)[2]), rep("P12 Inguinal", dim(merrick_p12)[2])), c(p21$capy_classification_merrick_etal, merrick_p12$merrick_class)))
skin_inguinal_celltypes$IDs_best <- skin_inguinal_celltypes$Var2
```

```{r}
skin_inguinal_celltypes$IDs_best <- gsub("adipocyte", "Adipocyte", skin_inguinal_celltypes$IDs_best)
skin_inguinal_celltypes$IDs_best <- gsub("committed_preAdipocyte", "Committed preadipocyte", skin_inguinal_celltypes$IDs_best)
skin_inguinal_celltypes$IDs_best <- gsub("endothelial", "Endothelial", skin_inguinal_celltypes$IDs_best)
skin_inguinal_celltypes$IDs_best <- gsub("group_3_cd142", "Group 3 (Cd142+)", skin_inguinal_celltypes$IDs_best)
skin_inguinal_celltypes$IDs_best <- gsub("group_4", "Group 4 (Wnt6+)", skin_inguinal_celltypes$IDs_best)
skin_inguinal_celltypes$IDs_best <- gsub("group_5", "Group 5", skin_inguinal_celltypes$IDs_best)
skin_inguinal_celltypes$IDs_best <- gsub("group_6", "Group 6", skin_inguinal_celltypes$IDs_best)
skin_inguinal_celltypes$IDs_best <- gsub("interstitial_progenitor", "Interstitial progenitor", skin_inguinal_celltypes$IDs_best)
skin_inguinal_celltypes$IDs_best <- gsub("Multi_ID", "Multi ID", skin_inguinal_celltypes$IDs_best)
skin_inguinal_celltypes$IDs_best <- gsub("neural_crest", "Neural crest", skin_inguinal_celltypes$IDs_best)
skin_inguinal_celltypes$IDs_best <- gsub("smooth_muscle", "Smooth muscle", skin_inguinal_celltypes$IDs_best)
```

```{r}
p1er <- ggplot(skin_inguinal_celltypes, aes(x = Var1, y = Freq, fill = IDs_best)) + geom_col(position = "fill", width = 0.9) + theme(axis.line.y = element_line(), axis.title.y = element_blank(), panel.background = element_blank(), strip.background = element_blank(), legend.position = "none", text = element_text(size = 8), plot.margin = margin(0, 0, 0, 0)) + scale_fill_manual(name = "Cell type", values = colpal12) + labs(x = "Tissue")
p1er
```

### Compare calls to consolidated reference

```{r}
merrick_compare <- data.frame(og = p21$capy_classification_merrick_etal, super = super_ref_class$call)
merrick_compare <- data.frame(table(merrick_compare$og, merrick_compare$super))
merrick_compare$Var1 <- as.character(merrick_compare$Var1)
merrick_compare$Var2 <- as.character(merrick_compare$Var2)
```

```{r}
merrick_compare[which(merrick_compare$Var1 == merrick_compare$Var2),]
```

```{r}
merrick_compare[which(merrick_compare$Var1 != merrick_compare$Var2), ]
```

```{r}
for(i in names(table(merrick_compare$Var1))) {
  print(paste0("cells lost from original ", i, " cell type: ", sum(merrick_compare$Freq[which(merrick_compare$Var1 == i)]) - merrick_compare$Freq[which(merrick_compare$Var1 == i & merrick_compare$Var2 == i)]))
}

```


```{r}
merrick_compare_mod <- merrick_compare[which(merrick_compare$Var1 == merrick_compare$Var2),]
merrick_compare_mod$Var2 <- NULL
merrick_compare_mod$instance <- "5 references"
merrick_compare_mod$Freq_norm <- merrick_compare_mod$Freq / sum(merrick_compare$Freq)
```


```{r}
merrick_plot <- data.frame(table(p21$capy_classification_merrick_etal))
merrick_plot$instance <- "Merrick reference only"
merrick_plot$Freq_norm <- merrick_plot$Freq / sum(merrick_plot$Freq)
merrick_plot <- rbind(merrick_plot, merrick_compare_mod)
```

```{r}
colpal12 <- c("#FFC33B", "#FF6E3A", "#E20134", "#9F0162", "#FFB2FD", "#00C2F9", "#008DF9", "#8400CD", "#00FCCF", "#FF5AAF", "#009F81", "gray")
```

```{r}
merrick_plot$instance <- factor(merrick_plot$instance, levels = c("Merrick reference only", "5 references"))
```

```{r}
ggplot(merrick_plot, aes(x = instance, y = Freq_norm, fill = Var1)) + geom_col() + theme_classic() + xlab("Instance") + ylab("Relative Abundance") + scale_fill_manual(name = "Cell type", values = colpal12)
```

```{r}
merrick_plot$Var1 <- as.character(merrick_plot$Var1)
merrick_plot <- rbind(merrick_plot, c("Other", 2201, "5 references", 2201/14478))
merrick_plot$Var1 <- factor(merrick_plot$Var1, levels = c("Other", "adipocyte", "committed_preadipocyte", "endothelial", "group_3_cd142", "group_4", "group_5", "group_6", "interstitial_progenitor", "Multi_ID", "neural_crest", "smooth_muscle", "Unassigned"))
merrick_plot$instance <- factor(merrick_plot$instance, levels = c("Merrick reference only", "5 references"))
merrick_plot$Freq_norm <- as.double(merrick_plot$Freq_norm)
```

```{r}
s1d <- ggplot(merrick_plot, aes(x = instance, y = Freq_norm, fill = Var1)) + geom_col() + theme_classic() + xlab("Instance") + ylab("Relative Abundance") + scale_fill_manual(name = "Cell type", values = c("darkgray", colpal12)) + theme(legend.position = "none")
s1d
```

### Hybrid cells

```{r}
s1f <- DimPlot(p21, cells.highlight = colnames(p21)[which(p21$capy_classification_merrick_etal == "Multi_ID")]) + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), axis.title = element_blank(), text = element_text(size = 8, family='Helvetica'), plot.margin = margin(0, 0, 0, 0), plot.title = element_blank(), legend.position = "none")
s1f
```

```{r}
hybrid_ids <- data.frame(adipocyte = rep(0, 10), 
                         committed_preadipocyte = rep(0, 10), 
                         endothelial = rep(0, 10), 
                         group_3_cd142 = rep(0, 10),
                         group_4 = rep(0, 10),
                         group_5 = rep(0, 10),
                         group_6 = rep(0, 10),
                         interstitial_progenitor = rep(0, 10),
                         neural_crest = rep(0, 10),
                         smooth_muscle = rep(0, 10))
rownames(hybrid_ids) <- colnames(hybrid_ids)
```

```{r}
hybrids_table <- data.frame(table(p21$capy_hybrid_merrick_etal))
```

```{r}
count <- 0
cellscount <- 0
for(i in 1:dim(hybrids_table)[1]) {
  types <- strsplit(as.character(hybrids_table$Var1[i]), "-")
  if(length(types[[1]]) == 2) {
    hybrid_ids[types[[1]][1], types[[1]][2]] <- hybrid_ids[types[[1]][1], types[[1]][2]] + hybrids_table$Freq[i]
  }
  else {
    count <- count + 1
    cellscount <- cellscount + hybrids_table$Freq[i]
  }
}
print(count)
print(cellscount)
hybrid_ids <- reshape2::melt(hybrid_ids)
hybrid_ids$variable2 <- rep(unique(hybrid_ids$variable), 10)
```

```{r}
ggplot(hybrid_ids, aes(x=variable, y=variable2, size=ifelse(value==0, NA,  value / sum(hybrid_ids$value)))) + geom_point(aes(color = variable)) +
  #scale_color_manual(values = colpal12[c(8, 4, 5, 6, 7, 2, 1)]) +
  scale_size_area(name = "Fraction of adipocyte lineage hybrids", max_size=6) +
  guides(fill = FALSE, color = FALSE) +
  theme(legend.position = "bottom",
        #text = element_text(size = 8),
        axis.ticks = element_blank(), 
        axis.title = element_blank(), 
        #title = element_text(size = 8),
        legend.key=element_blank(),
        legend.box.spacing = unit(0, "in"),
        legend.margin = margin(0, 0, 0, 0),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line.x = element_line(color = "black", size = 1),
        axis.line.y = element_line(color = "black", size = 1), axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
```


We subsetted hybrids to visualize only adipose cell type hybrids, as well as re-ordered this table for plotting. This file can be found on GitHub:

```{r}
hybrid_ids_reordered <- read.csv("./hybrid_ids_reordered_capy_merrick_for_plotting_fig1.csv")

hybrid_ids_reordered_sub <- hybrid_ids_reordered[hybrid_ids_reordered$variable %in% c("Interstitial progenitor", "Committed preadipocyte", "Group 3", "Adipocyte", "Group 5", "Group 4", "Group 6") & hybrid_ids_reordered$variable2 %in% c("Interstitial progenitor", "Committed preadipocyte", "Group 3", "Group 4", "Group 5", "Group 6"), ]
```

```{r}
hybrid_ids_reordered_sub$variable <- gsub("Interstitial progenitor", "Interstitial\nprogenitor", hybrid_ids_reordered_sub$variable)
hybrid_ids_reordered_sub$variable <- gsub("Committed preadipocyte", "Committed\npreadipocyte", hybrid_ids_reordered_sub$variable)
hybrid_ids_reordered_sub$variable <- factor(hybrid_ids_reordered_sub$variable, levels = c("Interstitial\nprogenitor", "Group 3", "Group 4", "Group 5", "Group 6", "Committed\npreadipocyte", "Adipocyte"))
hybrid_ids_reordered_sub$variable2 <- gsub("Interstitial progenitor", "Interstitial\nprogenitor", hybrid_ids_reordered_sub$variable2)
hybrid_ids_reordered_sub$variable2 <- gsub("Committed preadipocyte", "Committed\npreadipocyte", hybrid_ids_reordered_sub$variable2)
hybrid_ids_reordered_sub$variable2 <- factor(hybrid_ids_reordered_sub$variable2, levels = c("Interstitial\nprogenitor", "Group 3", "Group 4", "Group 5", "Group 6", "Committed\npreadipocyte", "Adipocyte"))
```


```{r}
s1g <- ggplot(hybrid_ids_reordered_sub, aes(x=variable, y=variable2, size=ifelse(value==0, NA,  value / sum(hybrid_ids_reordered_sub$value)))) + geom_point(aes(color = variable)) +
  scale_color_manual(values = colpal12[c(8, 4, 5, 6, 7, 2, 1)]) +
  scale_size_area(name = "Fraction of adipocyte lineage hybrids", max_size=6) +
  guides(fill = FALSE, color = FALSE) +
  theme(legend.position = "bottom",
        #text = element_text(size = 8),
        axis.ticks = element_blank(), 
        axis.title = element_blank(), 
        #title = element_text(size = 8),
        legend.key=element_blank(),
        legend.box.spacing = unit(0, "in"),
        legend.margin = margin(0, 0, 0, 0),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        axis.line.x = element_line(color = "black", size = 1),
        axis.line.y = element_line(color = "black", size = 1), axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
s1g
```

## New proposed classification strategy and visualization

```{r}
p21$new_classification <- NA
p21$new_classification[p21$seurat_clusters %in% c(0, 1, 2, 4, 5, 6, 10)] <- "Interstitial progenitor"
p21$new_classification[p21$seurat_clusters == 7] <- "Committed preadipocyte"
p21$new_classification[p21$seurat_clusters %in% c(3, 7, 8, 11) & p21$capy_classification_merrick_etal == "interstitial_progenitor"] <- "Transitioning interstitial progenitor"
p21$new_classification[colnames(preadips)[preadips$seurat_clusters %in% c(0, 1, 7, 8)]] <- "Immature preadipocyte"
```

```{r}
p21$new_classification_short <- gsub("Interstitial progenitor", "Prog", p21$new_classification)
p21$new_classification_short <- gsub("Transitioning interstitial progenitor", "Trans prog", p21$new_classification_short)
p21$new_classification_short <- gsub("Immature preadipocyte", "Imm preadip", p21$new_classification_short)
p21$new_classification_short <- gsub("Committed preadipocyte", "Comm preadip", p21$new_classification_short)
p21$new_classification <- factor(p21$new_classification, levels = c("Interstitial progenitor", "Transitioning interstitial progenitor", "Immature preadipocyte", "Committed preadipocyte"))
p21$new_classification_short <- factor(p21$new_classification_short, levels = c("Prog", "Trans prog", "Imm preadip", "Comm preadip"))
```

```{r}
p1f <- DimPlot(p21, group.by = "new_classification", cols = colpal6[1:4]) + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), axis.title = element_blank(), text = element_text(size = 8, family='Helvetica'), plot.margin = margin(0, 0, 0, 0), plot.title = element_blank(), legend.position = "bottom")
p1f
```



```{r}
p1gl <- VlnPlot(p21, features = "Dpp4", group.by = "new_classification_short", pt.size = 0, cols = colpal6[1:4]) + stat_summary(fun = mean, geom='point', size = 3, colour = "black") + NoLegend()
p1gl
p1gm <- VlnPlot(p21, features = "Cd9", group.by = "new_classification_short", pt.size = 0, cols = colpal6[1:4]) + stat_summary(fun = mean, geom='point', size = 3, colour = "black") + NoLegend()
p1gm
p1gr <- VlnPlot(p21, features = "F3", group.by = "new_classification_short", pt.size = 0, cols = colpal6[1:4]) + stat_summary(fun = mean, geom='point', size = 3, colour = "black") + NoLegend()
p1gr
```


```{r}
frxns_p21 <- p21@meta.data[, grep('frxn', colnames(p21@meta.data))]
```

```{r}
frxns_p21$cell_name <- rownames(p21@meta.data)
rownames(frxns_p21) <- frxns_p21$cell_name
frxns_p21$call <- p21$capy_classification_merrick_etal
frxns_p21 <- reshape2::melt(frxns_p21)
table(frxns_p21$value < 0)
frxns_p21$value[frxns_p21$value < 0] <- 0
```

```{r}
frxns_p21$value_norm <- apply(frxns_p21, 1, function(x) { as.double(x[4]) / sum(frxns_p21$value[frxns_p21$cell_name == x[1]]) })
order_new <- frxns_p21$cell_name[order(frxns_p21[frxns_p21$variable == "frxn_cell.type_interstitial_progenitor_merrick", "value_norm"], decreasing = T)]
frxns_p21$variable <- factor(frxns_p21$variable)
frxns_p21$cell_name <- factor(frxns_p21$cell_name, levels = order_new)

```

```{r}
s1it <- ggplot(frxns_p21, aes(x = cell_name, y = value_norm, fill = variable)) + geom_col(position = "fill") + scale_fill_manual(values = colpal12) + theme(legend.position = "none", axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank()) + labs(y = "Fractional cell type composition")
s1it
```

```{r}
s1im <- ggplot(frxns_p21, aes(x = cell_name, fill = call)) + geom_bar() + scale_fill_manual(values = colpal12) + theme(legend.position = "none", axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank())
s1im
```

```{r}
p21_new_class_order_new <- data.frame(cell_name = colnames(p21), new_call = p21$new_classification)
p21_new_class_order_new$cell_name <- factor(p21_new_class_order_new$cell_name, levels = order_new)
```

```{r}
s1ib <- ggplot(p21_new_class_order_new, aes(x = cell_name, fill = new_call)) + geom_bar() + scale_fill_manual(values = colpal6) + theme(legend.position = "none", axis.text.x = element_blank(), axis.ticks.x = element_blank(), axis.title.x = element_blank())
s1ib
```

```{r}
s1i <- grid.arrange(s1jt, s1jm, s1jb, nrow = 3, heights = unit(c(2, 0.5, 0.5), c("in", "in", "in")))
s1i
```

We have deposited our whole processed P21 dataset on Figshare. It has been analyzed according to the code in this notebook: 
https://figshare.com/s/396c5f4ebd3c23d20094

## Additional feature plots

```{r}
p1h1 <- FeaturePlot(p21, features = "Akr1c18", cols = c("lightgrey", "#211E71"), order = T) + NoLegend() + theme(line = element_blank(),
                                                                axis.title = element_blank(),
                                                                axis.text = element_blank(),
                                                                plot.title = element_blank(),
                                                                plot.margin = margin(0, 0, 0, 0))
p1h2 <- FeaturePlot(p21, features = "Smpd3", cols = c("lightgrey", "#211E71"), order = T) + NoLegend() + theme(line = element_blank(),
                                                                axis.title = element_blank(),
                                                                axis.text = element_blank(),
                                                                plot.title = element_blank(),
                                                                plot.margin = margin(0, 0, 0, 0))
p1h3 <- FeaturePlot(p21, features = "Gap43", cols = c("lightgrey", "#211E71"), order = T) + NoLegend() + theme(line = element_blank(),
                                                                axis.title = element_blank(),
                                                                axis.text = element_blank(),
                                                                plot.title = element_blank(),
                                                                plot.margin = margin(0, 0, 0, 0))
p1h4 <- FeaturePlot(p21, features = "Dpp4", cols = c("lightgrey", "#211E71"), order = T) + NoLegend() + theme(line = element_blank(),
                                                                axis.title = element_blank(),
                                                                axis.text = element_blank(),
                                                                plot.title = element_blank(),
                                                                plot.margin = margin(0, 0, 0, 0))
p1h1
p1h2
p1h3
p1h4
```

```{r}
p1h5 <- FeaturePlot(p21, features = "Mfap4", cols = c("lightgrey", "#741112"), order = T) + NoLegend() + theme(line = element_blank(),
                                                                axis.title = element_blank(),
                                                                axis.text = element_blank(),
                                                                plot.title = element_blank(),
                                                                plot.margin = margin(0, 0, 0, 0))
p1h6 <- FeaturePlot(p21, features = "Igfbp7", cols = c("lightgrey", "#741112"), order = T) + NoLegend() + theme(line = element_blank(),
                                                                axis.title = element_blank(),
                                                                axis.text = element_blank(),
                                                                plot.title = element_blank(),
                                                                plot.margin = margin(0, 0, 0, 0))
p1h7 <- FeaturePlot(p21, features = "Igf1", cols = c("lightgrey", "#741112"), order = T) + NoLegend() + theme(line = element_blank(),
                                                                axis.title = element_blank(),
                                                                axis.ticks = element_blank(),
                                                                axis.text = element_blank(),
                                                                plot.title = element_blank(),
                                                                plot.margin = margin(0, 0, 0, 0))
p1h8 <- FeaturePlot(p21, features = "Cd9", cols = c("lightgrey", "#741112"), order = T) + NoLegend() + theme(line = element_blank(),
                                                                axis.title = element_blank(),
                                                                axis.text = element_blank(),
                                                                plot.title = element_blank(),
                                                                plot.margin = margin(0, 0, 0, 0))
p1h5
p1h6
p1h7
p1h8
```

```{r}
s1al <- FeaturePlot(p21, features = "Fabp4")
s1ar <- FeaturePlot(p21, features = "Pparg")
s1al
s1ar
```

```{r}
s1b <- VlnPlot(p21, features = "F3", pt.size = 0) + NoLegend() + xlab("Cluster")
s1b
```

## Supp 1 panels:

```{r}
s1e <- VlnPlot(p21, features = "F3", pt.size = 0) + NoLegend() + theme(plot.title = element_blank())
s1e
```

```{r}
s1jl <- FeaturePlot(p21, features = "Fabp4", order = TRUE) + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), axis.title = element_blank(), text = element_text(size = 8, family='Helvetica'), plot.margin = margin(0, 0, 0, 0), plot.title = element_blank())
s1jl
```

```{r}
s1jr <- FeaturePlot(p21, features = "Pparg", order = TRUE) + theme(axis.text = element_blank(), axis.ticks = element_blank(), axis.line = element_blank(), axis.title = element_blank(), text = element_text(size = 8, family='Helvetica'), plot.margin = margin(0, 0, 0, 0), plot.title = element_blank())
s1jr
```









F3 fig 1 stats

```{r}
t.test(adi0906@assays$RNA@data["F3", which(adi0906$new_classification == "Interstitial progenitor")],
       adi0906@assays$RNA@data["F3", which(adi0906$new_classification == "Transitioning interstitial progenitor")])
```

```{r}
t.test(adi0906@assays$RNA@data["F3", which(adi0906$new_classification == "Immature preadipocyte")],
       adi0906@assays$RNA@data["F3", which(adi0906$new_classification == "Transitioning interstitial progenitor")])
```

```{r}
t.test(adi0906@assays$RNA@data["F3", which(adi0906$new_classification == "Immature preadipocyte")],
       adi0906@assays$RNA@data["F3", which(adi0906$new_classification == "Committed preadipocyte")])
```
